---
title: "Lab 11"
author: "Wiktor Soral"
date: "December 12th 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Linear regression with multiple predictors

## Working with many dimensions

```{r}
x <- rnorm(100, 5, 1)
y <- rnorm(100, 5, 1)
z <- 2 + 1.5*x + 2.5*y + rnorm(100, 0, 2)
library(plot3D)
scatter3D(x,y,z, theta=45, phi=10,col = 'black')
```

## Working with many dimesions

```{r}
xs <- seq(min(x),max(x),length.out = 10)
ys <- seq(min(y),max(y),length.out = 10)
M <- mesh(xs, ys)
zs <- 2+1.5*M$x + 2.5*M$y
scatter3D(x,y,z, theta=45, phi=10,col = 'black')
surf3D(M$x,M$y, zs,add = T)
```

## Working with many dimesions

```{r}
library(ggplot2)
df <- as.data.frame(cbind(x,y,z))
suppressWarnings(qplot(x,z,data=df, geom=c('point','smooth'),method='lm'))
```

## Working with many dimensions

```{r}
suppressWarnings(qplot(y,z,data=df, geom=c('point','smooth'),method='lm'))
```

## Bayesian linear model

- $y \sim \mathcal{N}(\mu, \sigma^2)$
- $\mu = \beta_0 + \sum_k \beta_k * x_k$
- in other words $y$ is distributed according to Normal distribution with mean $\mu$ and variance $\sigma^2$
- note that $\mu$ changes (is conditional) on the value of $x_k$
- where $x_k$ is the $kth$ predictor

## Interactions

- $\mu = \beta_0 + \beta_1x_1 + \beta_2x_2 + \beta_3x_1x_2$
- $\mu = (\beta_0 + \beta_1x_1) + (\beta_2 + \beta_3x_1)x_2$
- $\mu = (\beta_0 + \beta_2x_2) + (\beta_1 + \beta_3x_2)x_1$
- in other words with interaction term $x_1x_2$ slope of $x_2$ can be represented as conditional on the values of $x_1$
- similarly slope of $x_1$ can be represented as conditional on the values of $x_2$

