---
title: "Lab 13"
author: "Wiktor Soral"
date: "January 23rd 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
```

# Bayesian Analysis of Factorial Desings

## Why factorial designs?

- Say we are interested in how some miracle diet affects weight loss.
- We are conducting a simple study and compare no diet to miracle diet.

## Why factorial designs?

```{r}
diet <- gl(2, 10, length = 40, labels = c('no diet', 'miracle diet'))
workout <- gl(2, 20, labels = c('watching TV', 'best workout'))
mu <- 3
tau_diet <- 4
tau_wout <- 6
tau_int <- -14
X <- model.matrix(~diet*workout)
wl <- X %*% c(mu, tau_diet, tau_wout, tau_int)+rnorm(40,0,1)
df <- data.frame(diet,workout, wl)
qplot(diet, wl, data=df, geom='boxplot', xlab='',ylab='Weight loss')
```

## Why factorial designs?

- Ahh, the miracle diet...What is really important is a proper workout.
- We should conduct another experiment, should we?

## Why factorial designs?

```{r}
qplot(workout, wl, data=df, geom='boxplot', xlab='',ylab='Weight loss')
```

## Why factorial designs?

- If we know in advance important *n* factors responsible for changes in DV, we should rather conduct *n* factorial experiment.
- E.g. with diet and workout.

## Why factorial designs?

```{r}
qplot(workout, wl, data=df, geom='boxplot', xlab='',ylab='Weight loss',
      colour=diet)
```

## Model for factorial designs

$y_{ijk} = \alpha + \beta_{1j} + \beta_{2k} + \beta_{1\times2jk} + \epsilon_{ijk}$

- $\alpha$ - grand mean
- $\beta_{1j}$ - effect of factor 1 for level indicated by subscript $j$
- $\beta_{2k}$ - effect of factor 2 for level indicated by subscript $k$
- $\beta_{1\times2jk}$ - interactive effect of factors 1 and 2 for their levels indicated by subscripts $j$ and $k$
- $\epsilon_{ijk}$ - individual level variation for a subject $i$ in groups $j$ of factor 1 and $k$ of factor 2

- we assume that $E(\epsilon_{ijk})$ and $Var(\epsilon_{ijk}) = \sigma^2$, ie. variance does not depend on group membership

## Understanding the model

- Say $\alpha = 4.5$

|             | no diet  | diet  |
|-------------|----------|-------|
| watching TV |   4.5    |  4.5  |
| workout     |   4.5    |  4.5  |

## Understanding the model

- Say $\beta_{1j}$ refers to diet, and is equal to a vector of length equal to the number of levels of diet: (-2, 2)

|             | no diet  | diet  |
|-------------|----------|-------|
| watching TV |   2.5    |  6.5  |
| workout     |   2.5    |  6.5  |

## Understanding the model

- Say $\beta_{2j}$ refers to workout, and is equal to a vector of length equal to the number of levels of workout: (-3, 3)

|             | no diet  | diet  |
|-------------|----------|-------|
| watching TV |   1.5    |  1.5  |
| workout     |   7.5    |  7.5  |

## Understanding the model

- The additive effect of $\beta_{1j}$ and $\beta_{2j}$ is just a combination of previous operations


|             | no diet  | diet  |
|-------------|----------|-------|
| watching TV |  -0.5    |  3.5  |
| workout     |   5.5    |  9.5  |

## Understanding the model

- What's left are interaction terms, which are specific to each cell (ie. each combination of factors)
- Say we have this interaction effects:


|             | no diet  | diet  |
|-------------|----------|-------|
| watching TV |  -1      |   1   |
| workout     |   1      |  -1   |

## Understanding the model

- When we add respective cells we will obtain estimated means

|             | no diet  | diet  |
|-------------|----------|-------|
| watching TV |  -1.5    |  4.5  |
| workout     |   6.5    |  8.5  |

## Bayesian factorial model

- The implementation follows the guidelines of Kruschke
- We assume $y_{ijk} \sim \mathcal{N}(\mu_{jk}, \sigma^2_y)$
- $\sigma^2_y \sim \mathcal{U}(low,high)$ where low and high indicate values respective to the scale of the DV
- $\mu_{jk} = \alpha + \beta_{1j} + \beta_{2k} + \beta_{1\times2jk}$
- $\alpha \sim \mathcal{N}(M, S^2)$ where M and S are mean and multiple of observed DV variance
- $\beta_{xj} \sim \mathcal{N}( 0.0, \sigma^2_x )$, are deflections related to $x$ (where $x$ is factor 1, 2, or interaction)
- $\sigma^2_x \sim \mathcal{G}(Shape,Rate)$, are variances/standard deviations of deflections related to $x$ (where $x$ is factor 1, 2, or interaction), and shape and rate are calculated w.r.t. distribution of DV 