---
title: "Lab 4"
author: "Wiktor Soral"
date: "October 24 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Working with R

## R

- free and open-source
- highly popular both in academia and business
- powerful, flexible, and capable
- works on almost every OS, both on desktop computers and in cloud
- integrates well with other software (e.g. SPSS)
- **powerful packages for Bayesian analysis (MCMCpack, rjags, rstan)**

## R plotting capabilities

```{r}
y1 <- round(rnorm(n = 36, mean = 7, sd = 2)) # Simulate data from normal distribution
y2 <- round(rnorm(n = 36, mean = 21, sd = 6))
y3 <- round(rnorm(n = 36, mean = 50, sd = 8))
x <- rep(LETTERS[1:12], 3)
grp <- rep(c("Grp 1", "Grp 2", "Grp 3"), each = 12)
dat <- data.frame(grp, x, y1, y2, y3)

library(ggplot2)
# Using default theme

kobe_theme2 <- function() {
  theme(
    legend.position = "bottom", legend.title = element_text(family = "Impact", colour = "#552683", size = 10),
    legend.background = element_rect(fill = "#E2E2E3"),
    legend.key = element_rect(fill = "#E2E2E3", colour = "#E2E2E3"),
    legend.text = element_text(family = "Impact", colour = "#E7A922", size = 10),
    plot.background = element_rect(fill = "#E2E2E3", colour = "#E2E2E3"),
    panel.background = element_rect(fill = "#E2E2E3"),
    axis.text = element_text(colour = "#E7A922", family = "Impact"),
    plot.title = element_text(colour = "#552683", face = "bold", size = 18, vjust = 1, family = "Impact"),
    axis.title = element_text(colour = "#552683", face = "bold", size = 13, family = "Impact"),
    panel.grid.major.y = element_line(colour = "#E7A922"),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    strip.text = element_text(family = "Impact", colour = "white"),
    strip.background = element_rect(fill = "#E7A922"),
    axis.ticks = element_line(colour = "#E7A922")
  )
}

x_id <- rep(12:1, 3) # use this index for reordering the x ticks
p3 <- ggplot(data = dat, aes(x = reorder(x, rep(1:12, 3)), y = y3, group = factor(grp))) +
  geom_bar(stat = "identity", fill = "#552683") + coord_polar() + facet_grid(. ~ grp) +
  ylab("Y LABEL") + xlab("X LABEL") + ggtitle("TITLE OF THE FIGURE")
p3 + kobe_theme2()
```

## R plotting capabilities

```{r}
p2 <- ggplot(data = dat, aes(x = x, y = y2, group = factor(grp))) +
  geom_line(stat = "identity", aes(linetype = factor(grp)), size = 0.7, colour = "#552683") +
  ylab("Y LABEL") + xlab("X LABEL") + ggtitle("TITLE OF THE FIGURE")
p2 + kobe_theme2() + scale_linetype_discrete("GROUP")
```

## R plotting capabilities

```{r}


kobe_theme <- function() {
  theme(
    plot.background = element_rect(fill = "#E2E2E3", colour = "#E2E2E3"),
    panel.background = element_rect(fill = "#E2E2E3"),
    axis.text = element_text(colour = "#E7A922", family = "Impact"),
    plot.title = element_text(colour = "#552683", face = "bold", size = 18, vjust = 1, family = "Impact"),
    axis.title = element_text(colour = "#552683", face = "bold", size = 13, family = "Impact"),
    panel.grid.major.x = element_line(colour = "#E7A922"),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    strip.text = element_text(family = "Impact", colour = "white"),
    strip.background = element_rect(fill = "#E7A922"),
    axis.ticks = element_line(colour = "#E7A922")
  )
}
p1 <- ggplot(data = dat, aes(x = reorder(x, x_id), y = y1)) + geom_bar(stat = "identity", fill = "#552683") +
  coord_flip() + ylab("Y LABEL") + xlab("X LABEL") + facet_grid(. ~ grp) +
  ggtitle("TITLE OF THE FIGURE")
p1 + kobe_theme()
```

## Steep learning curve

- intimidating for beginners
- challenging for intermediate users
- however with proper approach it can be quite fun
- forget that R is a programming language
- and think about it as a statistical package (like SPSS)

## Some basic advices

- know where to find the newest version R <a href="https://cran.r-project.org">cran.r-project.org</a>
- use RStudio <a href="www.rstudio.com">www.rstudio.com</a> - the more advanced you become the more useful it will be
- know how to find help <a href="www.statmethods.net">www.statmethods.net</a> - read the docs!
- keep yourself engaged and be a part of the community: <a href="www.r-bloggers.com">www.r-bloggers.com</a>

## How not to loose your mind

- **always** save your code
- keep your code clear and organized: use comments, indentation, and spacing
- at least skim through Help pages before running a new function (e.g. ?lm)
- try do some working examples and understand how they work
- **do not** save your global environment when exiting R
- setting your working directory (setwd) might be a good idea (instead of working in a root directory)


## Errors happen 

- don't panic - know how to distinguish messages, warnings, and errors
- carefully read the error message
- check your syntax, usually it is just a misspelled name, remember that R is case-sensitive (Rstudio **is not** RStudio)*
- read the docs - check the Usage and Arguments sections
- try to localize the place that causes the error and see what you can do
- *google* the error message





