---
title: "Lab 5"
author: "Wiktor Soral"
date: "October 31 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Monte Carlo simulations

## Calculating evidence {.flexbox .vcenter}

$P(\theta | \mathcal{D}) = \frac{P(\mathcal{D} | \theta) \times P(\theta)}{P(\mathcal{D})}$

posterior = likelihood $\times$ prior / evidence

How to calculate evidence?

## Calculating evidence {.flexbox .vcenter}

With only 2 or 3 values of $\theta$ it is fairly simple (recall HW 1):

$P(\mathcal{D}) = [P(\mathcal{D} | \theta = 1) \times P(\theta = 1)] + [P(\mathcal{D} | \theta = 2) \times P(\theta = 2)]$

In general we could write:

$P(\mathcal{D}) = \sum_{\theta} P(\mathcal{D} | \theta) \times P(\theta)$

## Calculating evidence {.flexbox .vcenter}

Though, usually we will use $\theta$ with continuous scale (i.e. infinite number of values).

In math we denote summing across values of continuous scale as **integrating**:

$P(\mathcal{D}) = \int_{\theta} P(\mathcal{D} | \theta) \times P(\theta) \mathrm{d\theta}$

## Calculating integrals - analytical app.

```{r}
library(ggplot2)
curve(x^2, from = 0, to=4, lwd=3, col='red', ylab=expression(x^2),
      main='Integral is a function used to calculate the area below the curve')
x=c(0,seq(0, 4, length.out = 50), 4, 0)
y=c(0,seq(0, 4, length.out = 50)^2, 0, 0)
polygon(x,y, border = 'pink', col=alpha(colour = 'red', alpha = 0.2))
```

We know from analytical solution: $\int x^2 \mathrm{dx} = \frac{x^3}{3} + const$

## Calculating integrals - numerical app.

```{r}
bins=9
xs = seq(0,4, length=bins)[1:(bins-1)]
step = xs[2] - xs[1]
curve(x^2, from = 0, to=4, lwd=3, col='red', ylab=expression(x^2),
      main='Riemann integral (8 bins) - sum of boxes = 17.5')
rects = c()
for (i in xs){
  rect(i, 0, i+step, i^2, border = 'pink', col=alpha(colour = 'red', alpha = 0.2))
  sRect = step * (i^2)
  rects = c(rects, sRect)
}
```

## Calculating integrals - numerical app.

```{r}
bins=17
xs = seq(0,4, length=bins)[1:(bins-1)]
step = xs[2] - xs[1]
curve(x^2, from = 0, to=4, lwd=3, col='red', ylab=expression(x^2),
      main='Riemann integral (16 bins) - sum of boxes = 19.375')
rects = c()
for (i in xs){
  rect(i, 0, i+step, i^2, border = 'pink', col=alpha(colour = 'red', alpha = 0.2))
  sRect = step * (i^2)
  rects = c(rects, sRect)
}
```

## Calculating integrals - numerical app.

```{r}
bins=33
xs = seq(0,4, length=bins)[1:(bins-1)]
step = xs[2] - xs[1]
curve(x^2, from = 0, to=4, lwd=3, col='red', ylab=expression(x^2),
      main='Riemann integral (32 bins) - sum of boxes = 20.343')
rects = c()
for (i in xs){
  rect(i, 0, i+step, i^2, border = 'pink', col=alpha(colour = 'red', alpha = 0.2))
  sRect = step * (i^2)
  rects = c(rects, sRect)
}
```

## Calculating integrals - numerical app.

```{r}
bins=65
xs = seq(0,4, length=bins)[1:(bins-1)]
step = xs[2] - xs[1]
curve(x^2, from = 0, to=4, lwd=3, col='red', ylab=expression(x^2),
      main='Riemann integral (64 bins) - sum of boxes = 20.836')
rects = c()
for (i in xs){
  rect(i, 0, i+step, i^2, border = 'pink', col=alpha(colour = 'red', alpha = 0.2))
  sRect = step * (i^2)
  rects = c(rects, sRect)
}
```

## Computing integrals - numerical app.

- With 1 parameter/dimension CPU has to perform $100$ operations
- With 2 parameters/dimensions CPU has to perform $100^2 = 10000$ operations
- With 3 parameters/dimensions CPU has to perform $100^3 = 1000000$ operations

- With 10 parameters/dimensions CPU has to perform $100^{10} = 100000000000000000000$ operations

- Numerical approach quickly becomes too hard even for fast computers

## Sampling

```{r}
curve(x^2, from = 0, to=4, lwd=3, col='red', ylim=c(0, 17), ylab=expression(x^2),
      main='Rejection sampling - pink rect area = 68')
rect(0, 0, 4, 17, border='pink', lwd=2)
```

## Sampling

```{r}
curve(x^2, from = 0, to=4, lwd=3, col='red', ylim=c(0, 17), ylab=expression(x^2),
      main='Rejection sampling - pink rect area = 68')
rect(0, 0, 4, 17, border='pink', lwd=2)
xs = runif(1000, 0, 4)
ys = runif(1000, 0, 17)
points(xs, ys, col='pink')
```

## Sampling

```{r}
curve(x^2, from = 0, to=4, lwd=3, col='red', ylim=c(0, 17), ylab=expression(x^2),
      main='Rejection sampling - pink rect area = 68')
rect(0, 0, 4, 17, border='pink', lwd=2)
xs_rej = xs[ys <= xs^2]
ys_rej = ys[ys <= xs^2]
points(xs, ys, col='pink')
points(xs_rej, ys_rej, col='pink', pch=16)
```

E.g. 318 of 1000 sampled points are below the curve - 0.318. $68 \times 0.318 = 21.624$