---
title: "Lab 8"
author: "Wiktor Soral"
date: "November 21st 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Bayesian Estimation Supersedes T-test

## Rationale

- Real-life data rarely concord with assumptions of common statistical approaches
- Group variances usually differ (whereas with t-test we assume they are equal)
- Tails of the sampled distribution are usually heavy and include outliers (whereas we assume the data follow Normal distribution)
- Finding proper transformation is often difficult, and removing outliers is usually controversial

## Rationale

- Instead of 'correcting' our data we can try to model them as they are
- JAGS gives us enough flexibility to deal with most of the real-life distributional issues
- Make sure you to read chapter 16 from DBDA
- If you would like to include citation to the method we will cover today go to: <a href="http://psycnet.apa.org/record/2012-18082-001">http://psycnet.apa.org/record/2012-18082-001</a>

## t distribution

Using t-distribution (instead of Normal) is a way to obtain robust (to outliers) estimates of group variances

```{r}
curve(dt(x, 30), from=-5, to=5, lwd=2, col='lightblue', ylab="",xlab="t distribution")
curve(dt(x, 4), from=-5, to=5, lwd=2, lty=2, col='lightblue', add=TRUE)
curve(dt(x, 1), from=-5, to=5, lwd=2, lty=3, col='lightblue', add=TRUE)
legend('topright', legend=c(expression(nu == 30), expression(nu == 4),
                            expression(nu == 1)), lwd=c(2,2,2),
       col="lightblue", lty=1:3)
```

## Prior on $\nu$ parameter

```{r}
curve(dexp(x, 1/29), from=1, to=100, xlab=expression(nu),ylab="",
        type="h", lwd=2, col="lightblue", main="Exp(1/29)")
lines(c(29.75, 29.75), c(0,dexp(30, 1/29)), lwd=2, col="blue")
text(x = 30, y=0.015, labels = "Mean of the distribution = 30",
     adj=0)
```

## ROPE

- A region of practical equivalence (ROPE) indicates a small range of parameter values that are considered to be practically equivalent to the null value for purposes of the particular application.
- E.g. if we are assessing the efficacy of a drug versus a placebo, we might only consider using the drug if it improves the probability of cure by at least 5 percentage points. Thus, the ROPE on the diference of cure probabilities could have limits of ±0.05.
